The Units of Permissionless Consensus: \\ Towards Mobile and Edge Computing

### 2022-11-16 theoretical-consensus/article/article.tex
add page 09

\begin{abstract}
Disclaimer: This paper is a work in progress. It is not a final version and it is not intended to be submitted yet to the Distributed Systems Seminar. The content of this paper is subject to change.
\end{abstract}

\begin{IEEEkeywords}
IEEE, IEEEtran, journal, \LaTeX, paper, template.
\end{IEEEkeywords}

@@

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec auctor, nisl eget aliquam aliquam.

@@

This prototype was developed and shipped as a static web page.
The base layer tooling choice resided in the \emph{Nuxt.js} framework, a meta-framework
for building web applications based on \emph{Vue.js}. With that established,
it became clear the need for a protocol for syncing and communicating in a 
peer-to-peer way between the devices. The choice resided in the \emph{GUN} ecosystem,
which ships a small but powerful set of tools for in-browser data synchronization.
Since there was \emph{TypeScript} support from all the major dependencies of this project,
this was the language chosen for the interactive web development. The cryptographic standards used in the 
\emph{PoW} algorithm were imported from the \emph{crypto-js} library. And, finally, HTML and CSS, with the help
of the \emph{Tailwind CSS} framework were used for styling and presentation.

A major requirement for such a system is the capacity for parallel computations, 
in order to be able to, simultaneously, listen for and validate proposed blocks, 
present the current results in the screen, and mine new blocks. The traditional 
\emph{JavaScript} computation model is a single-threaded environment that, naturally,
does not allow for the efficient and effective handling of all these scenarios, happening
at the same time. Most of the modern web browsers have a living solution for this problem, 
and that is based on \emph{Web Workers}. This \emph{API} was used in order to offload
the expensive mining computations and take that burden out of the main thread, leaving it
with the responsibility of listening for information updates, either from other peers, with
proposed blocks, or from the \emph{miner} threads, with new mined blocks, and decide and update
the user interface accordingly.

@@

The prototype was tested, is demonstrable and works as intended. Some \emph{Nakamoto} chain-based
features were simplified, for the sake of maintaining a didactic complexity. For example, 
the Longest-Chain rule, in this prototype, works by having a temporary set of active blocks 
that get appended to the main chain if there are new blocks mined on top
of them. This setting enforces higher synchrony between the peers, as there is no mechanism for
general chain data synchronization. This also means that a new peer joining the network does not
need or does not ask for the rest of the chain history to start mining, but just the most recent block.
Again, this is justified by the purpose of this prototype which was to demonstrate the achievement
of consensus in a permissionless environment.

@@

>>>>

\begin{abstract}

This paper presents an overview of the consensus problem
applicable to permissionless environments. Starting from 
the classical roots of consensus in distributed systems,
and then transitioning to the more recent developments in the
field of permissionless consensus, this paper also reviews
the common approaches and architectural design of the most well known protocols, 
and discusses the trade-offs between the different approaches.
Finally, this paper points out the challenges that
permissionless consensus faces in the context of resource
constrained environments, such as mobile and edge computing,
identifies and argues about the current approaches to address these challenges,
and shows potential research directions.

\end{abstract}

\begin{IEEEkeywords}
Consensus, Distributed Systems, Permissionless Consensus, Mobile and Edge Computing, Blockchain, Resource Constrained Environments.
\end{IEEEkeywords}

@@

In this paper, we have presented a comprehensive summary of the permissionless
consensus problem, the most widely used consensus protocols, and the challenges
that arise when applying these protocols to resource-constrained networks.
We have presented the challenges that these networks are facing, and the 
limitations of the existing consensus protocols.
We have also presented the main approaches that have been explored to address these challenges,
and we have showed potential paths for future research.

@@

This prototype was developed and shipped as a static web page.
The base layer tooling choice resided in the \emph{Nuxt.js} framework \cite{nuxt}, a meta-framework
for building web applications based on \emph{Vue.js} \cite{vuejs}. With that established,
it became clear the need for a protocol for syncing and communicating in a 
peer-to-peer way between the devices. The choice resided in the \emph{GUN} ecosystem \cite{gundb},
which ships a small but powerful set of tools for in-browser data synchronization.
Since there was \emph{TypeScript} support from all the major dependencies of this project,
this was the language chosen for the interactive web development \cite{typescript}. The cryptographic standards used in the 
\emph{PoW} algorithm were imported from the \emph{crypto-js} library \cite{crypto-js}. And, finally, HTML and CSS, with the help
of the \emph{Tailwind CSS} framework \cite{tailwindcss} were used for styling and presentation.

A major requirement for such a system is the capacity for parallel computations, 
in order to be able to, simultaneously, listen for and validate proposed blocks, 
present the current results in the screen, and mine new blocks. The traditional 
\emph{JavaScript} computation model is a single-threaded environment that, naturally,
does not allow for the efficient and effective handling of all these scenarios, happening
at the same time. Most of the modern web browsers have a living solution for this problem, 
and that is based on \emph{Web Workers} \cite{webworkers}. This \emph{API} was used in order to offload
the expensive mining computations and take that burden out of the main thread, leaving it
with the responsibility of listening for information updates, either from other peers, with
proposed blocks, or from the \emph{miner} threads, with new mined blocks, and decide and update
the user interface accordingly.

@@

The prototype was tested, is demonstrable and works as intended. Some \emph{Nakamoto} chain-based
features were simplified, for the sake of maintaining a didactic complexity. For example, 
the Longest-Chain rule, in this prototype, works by having a temporary set of active blocks 
that get appended to the main chain if there are new blocks mined on top
of them. This setting enforces higher synchrony between the peers, as there is also no mechanism for
general chain data synchronization. Additionally, this means that a new peer joining the network does not
need or does not ask for the rest of the chain history to start mining, but just the most recent block.
The difficulty was also set to a constant value, for the sake of simplicity.
Again, this is justified by the purpose of this prototype, which was to demonstrate the achievement
of consensus in a permissionless environment.

@@

The prototype was developed as a proof-of-concept and a demonstration of the feasibility
of the \emph{PoW} consensus protocol in a permissionless environment. The next steps
for this project would be to extend the functionality and the features of the system,
in order to make it more robust and usable. Some of the ideas for future work are, for example,
the improvement of the \emph{PoW} algorithm, by adding a difficulty adjustment mechanism,
to make the mining process more fair for computationally weaker devices, or to
add a mechanism for the synchronization of the old chain history. Another idea would be to
improve the parallelization of the computations, by more efficiently distributing the
workload among mining threads. Finally, the user interface could be improved 
with more informative and interactive features, that would allow the user to have a 
deeper understanding of the system and the consensus process.
